---
title: "Identify Counties using Zip"
date: '`r strftime(Sys.time(), format = "%B %d, %Y")`'
output: 
    html_document:
        toc: true
        toc_float: true
        code_folding: show
params:
    state_name: Xxxxxx # e.g., Missouri, Georgia, etc.
    db_file: "~/data2/XX_Dashboard/lic.sqlite3"
    csv_file: "~/data/XX_Dashboard/geocode-addresses/cust-for-geocode.csv"
    crosswalk_file: "~/Data2/Zip_County_Crosswalk/Zipcodes.csv"
    test: TRUE # If TRUE, won't write to sqlite database
---

```{r setup, message=FALSE}
library(tidyverse)
library(RSQLite)
library(tibble)
knitr::opts_chunk$set(comment = NA)
```

## Params

- This attaches a county fip to the standardized customer table
- Note: Geocoding is a preferable way to identify zipcode. The zip-to-county is used (for now) because we haven't implemented a reliable way to geocode millions of records.

```{r}
tribble(
    ~param, ~value,
    "state_name", params$state_name,
    "db_file", params$db_file,
    "csv_file", params$csv_file, 
    "crosswalk_file", params$crosswalk_file,
    "test", params$test
) %>% DT::datatable()
```

## Load Data

### Zip to County Relation

- Tom sent this to Dan on 2/10/14. Not sure about data source or processing.

```{r}
col_types <- cols(
  ID = col_skip(),
  ZIPCODE = col_character(),
  STATE_NAME = col_character(),
  STATE_FIPS = col_skip(),
  COUNTY_NAME = col_character(),
  COUNTY_FIPS = col_double(),
  Field6 = col_skip()
)
zip_cnty_in <- read_csv(params$crosswalk_file, col_types = col_types)
names(zip_cnty_in) <- c("zip", "state", "county", "fips")

zip_cnty <- filter(zip_cnty_in, state == params$state_name) %>% 
    select(-state) %>%
    data.frame()
# count(zip_cnty, zip)
# count(zip_cnty, fips)
glimpse(zip_cnty)
```

### Customer Zipcodes

- Columns may vary by state

```{r}
col_types <- cols(
  cust_id = col_integer(),
  street = col_skip(),
  city = col_skip(),
  state = col_skip(),
  zip = col_character()
)
cust_zips <- read_csv(params$csv_file, col_types = col_types)
# count(zips, zip)
glimpse(cust_zips)
```

### Cleaned Customer Data

```{r}
db <- src_sqlite(params$db_file)
cust <- tbl(db, "cust") %>% collect(n = Inf)
glimpse(cust)
```

## Attach County

### Resident Customers

```{r}
fips <- left_join(cust_zips, zip_cnty) %>% select(-zip)
count(fips, fips, county) %>% DT::datatable()

# check missing
missing_fips <- filter(fips, is.na(fips)) %>% nrow()
tribble(
    ~all_resident_customers, ~missing_fips, ~pct_missing,
    nrow(fips), missing_fips, missing_fips / nrow(fips) * 100
)
```

### All Customers

```{r}
cust <- left_join(cust, fips, by = "cust_id") %>%
    mutate(fips = as.integer(fips))
count(cust, res, fips, county) %>% DT::datatable()
cust <- select(cust, -county)
```

## Write to SQLite

- The only change should be one additional column (fips)

```{r}
glimpse(cust)
```

```{r}
if (!params$test) {
    con <- dbConnect(SQLite(), dbname = params$db_file)

    # create a new customer table (with zip) to insure data isn't lost by mistake
    # it would probably be more efficient to add a new column, but this works
    cust1 <- cust
    copy_to(db, cust1, temporary = FALSE)
    
    # remove original customer table and rename the copied table
    dbRemoveTable(con, "cust")
    dbGetQuery(con, "ALTER TABLE cust1 RENAME TO cust")
    dbDisconnect(con)
}
```

