---
title: "Create privilege tables by priv (hunt, fish, etc.)"
date: '`r strftime(Sys.time(), format = "%B %d, %Y")`'
output: 
    html_document:
        toc: true
        toc_float: true
        code_folding: hide
params:
    priv_nm: hunt  # (fish, hunt, all_sports) or (deer, trout, etc.)
    priv_hist: NONE  # (NONE, fish, hunt, etc.) > for subtype permissions (otherwise "NONE")
    lic_filter: type %in% c("hunt", "trap", "combo")
    yrs: 2007:2016   # (first_year:last_year)
    test: TRUE  # (TRUE, FALSE) > if TRUE, runs a sample & doesn't write output to sqlite
    db_file: "~/data2/XX_Dashboard/lic.sqlite3"  # sqlite db location (varies by state)
---

```{r setup, message=FALSE}
library(tidyverse)
library(RSQLite)
library(salic)
library(DT)
library(stringr)
knitr::opts_chunk$set(comment = NA)
```

## Params

```{r}
# make a vector of years from the params$yrs input
get_range <- function(x) {
    y <- stringr::str_split_fixed(x, ":", 2)
    `:`(y[1], y[2])
}
yrs <- get_range(params$yrs)
```

```{r}
tibble::tribble(
    ~param, ~value,
    "priv_nm", params$priv_nm,
    "lic_filter", params$lic_filter,
    "priv_hist", params$priv_hist,
    "yrs", params$yrs,
    "test", params$test,
    "db_file", params$db_file
) %>% DT::datatable()
```

## Load

### License Types

```{r}
db <- src_sqlite(params$db_file)
lic_all <- tbl(db, "lic") %>% collect()
lic <- filter_(lic_all, .dots = params$lic_filter)
DT::datatable(lic)
```

### Sales

```{r}
# get sales for relevant license types
# the if conidtion just insures the logic works on the sqlite backend (== vs %in%)
if (nrow(lic) == 1) {
   sale <- tbl(db, "sale") %>% filter(lic_id == lic$lic_id) 
} else {
    sale <- tbl(db, "sale") %>% filter(lic_id %in% lic$lic_id)
}
sale <- select(sale, cust_id, lic_id, year, res) %>% 
    filter(year %in% yrs) %>%
    collect(n = Inf)
glimpse(sale)
```

### Sample

- For testing, take a sample of about 10,000 license holders

```{r}
if (params$test) {
    cust <- distinct(sale, cust_id)
    sample_size <- min(10000, nrow(cust))
    cust <- sample_n(cust, sample_size)
    sale <- semi_join(sale, cust, by = "cust_id")
    rm(cust)
    glimpse(sale)
}
```

## Prepare Sales Data

### Check Duration

- State-specific coding may be required if duration is not completely accounted for in the license table

```{r}
count(lic, duration)
```

### Rank Sales

- Creates one row per customer-year (grouping by cust_id, year)
    + picks highest duration (e.g., a 5-year license would be chosen over a 1-year license)
    + picks Resident (res == 1) over Nonresident (res == 0). A customer generally shouldn't have 2 different residencies in the same year, but it happens occasionally

```{r}
sale_ranked <- sale %>%
    rank_sale(rank_var = c("duration", "res"), grp_var = c("cust_id", "year"))

# visual comparison
ranked_sale_count <- count(sale_ranked, res, year, duration) %>% mutate(grp = "ranked-sales")
sale_count <- count(sale, res, year, duration) %>% mutate(grp = "sales")
bind_rows(sale_count, ranked_sale_count) %>%
    ggplot(aes(factor(year), n, fill = factor(duration))) +
    geom_bar(stat = "identity") +
    facet_grid(factor_res(res) ~ grp, scales = "free_y")
```

## Prepare Privilege Data

### Make License History

- Note that each of the 3 steps produces an output for checking
    + it might be helpful to run these one at a time during testing
    + the data format is described in _Analyst Docs > Data Dictionary > license history
    
```{r, eval = params$priv_hist == "NONE", message = FALSE}
# if not in test mode then this can take a while to run

# 1. make license history
lic_history <- sale_ranked %>%
    select(cust_id, duration, year, res) %>%
    make_lic_history(yrs, carry_vars = "res")

# Visual Compare privs and sales
sale_count <- count(sale_ranked, res, year, duration) %>% 
    mutate(duration_run = duration, grp = "sales")
priv_count <- count(lic_history, res, year, duration_run) %>% 
    mutate(grp = "privs")
bind_rows(sale_count, priv_count) %>%
    ggplot(aes(factor(year), n, fill = factor(duration_run))) +
    geom_bar(stat = "identity") +
    facet_grid(factor_res(res) ~ grp, scales = "free_y")


# 2. Identify R3
# this condition means R3 won't be calcualated unless there is a sufficient license history
if (length(yrs) > 5) {
    lic_history <- identify_R3(lic_history, yrs)
    
    # visual check
    count(lic_history, res, year, R3) %>%
        ggplot(aes(factor(year), n, fill = factor_R3(R3))) +
        geom_bar(stat = "identity") +
        facet_wrap(~ factor_res(res), scales = "free_y", nrow = 2)

} else {
    lic_history$R3 <- NA 
}
check_identify_R3(lic_history, yrs) %>% DT::datatable()


# 3. Identify lapse
lic_history <- identify_lapse(lic_history, yrs)

# check
count(lic_history, res, year, lapse) %>%
    ggplot(aes(factor(year), n, fill = factor(lapse))) +
    geom_bar(stat = "identity") +
    facet_wrap(~ factor_res(res), scales = "free_y", nrow = 2)

check_identify_lapse(lic_history) %>% DT::datatable()


# 4. Overall Check
# This samples a few customers, showing relevant calculation variables
lic_history %>%
    select(cust_id, year, res, bought, duration, duration_run, lag_duration_run, R3, lapse) %>%
    check_history_samp() %>% DT::datatable()


# 5. Make priv table
# this drops the temporary cacluation variables (duration_run, etc.)
priv <- lic_history %>%
    select(cust_id, year, lapse, R3, res)
```

```{r, eval = params$priv_hist != "NONE", message = FALSE}
# get history for a specified reference privilege

priv_hist_nm <- paste("priv", params$priv_hist, sep = "_")
priv_hist <- tbl(db, priv_hist_nm) %>% 
    select(cust_id, year, lapse, R3) %>%
    collect(n = Inf)

# identify R3/lapse (by year) depending on that privilege
priv <- select_(sale_ranked, .dots = c("cust_id", "year", "res")) %>%
    inner_join(priv_hist, by = c("cust_id", "year"))
# todo: maybe add a warning here if nrow(priv) < nrow(priv_sale)
```

### Check Final Priv

-  This should show TRUE (i.e., all relevant sales are included)

```{r}
nrow(sale_ranked) == semi_join(priv, sale_ranked, by = c("cust_id", "year")) %>% nrow()
```

## Write to SQLite

### License Privilege Table

- The idea here is to have a separate table that explicitly identifies all the license types that go into a specific privilege
    + This allows a simple join to associate license types with privileges

```{r}
if (!(params$test)) {
    # match priv name with output table name
    # remove spaces in type name to insure compatibility with sqlite table names
    priv_nm <- stringr::str_replace_all(params$priv_nm, " ", "_")
    out_nm <- paste("priv", priv_nm, sep = "_")
    
    lic_priv <- select(lic, lic_id)
    lic_priv$priv <- out_nm
    lic_priv <- select(lic_priv, priv, lic_id)
    glimpse(lic_priv)
}
```


### Write Output Data

```{r}
if (!(params$test)) {
    con <- dbConnect(SQLite(), dbname = params$db_file)
    
    ### Privilege Table
    priv <- priv %>% mutate(
        year = as.integer(year), 
        lapse = as.integer(lapse), 
        R3 = as.integer(R3) 
    )
    glimpse(priv)
    
    # write table (overwrite if this privilege table already exists)
    if (out_nm %in% dbListTables(con)) dbRemoveTable(con, out_nm)
    dbWriteTable(con, out_nm, data.frame(priv))
    
    ### Lic Priv Table
    # create table if it doesn't yet exist
    if (!("lic_priv" %in% dbListTables(con))) {
        dbWriteTable(con, "lic_priv", data.frame(lic_priv))
    } else {
        # overwrite selected priv records to insure only the newest is kept
        lic_priv_old <- tbl(db, "lic_priv") %>% collect()
        lic_priv_old <- filter(lic_priv_old, priv != out_nm)
        lic_priv <- bind_rows(lic_priv, lic_priv_old)
        dbRemoveTable(con, "lic_priv")
        dbWriteTable(con, "lic_priv", data.frame(lic_priv))
    }
    dbDisconnect(con)
}
```

```{r}
sessionInfo()
```

