---
output: github_document
---
<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# salic

The salic R package lays out a workflow to prepare and summarize agency data. In particular, it aims to simplify the production of national/regional dashboards with a set of functions that work across state agencies.

## Installation

- First, install R: https://www.r-project.org/
- Then install salic and it's R package dependencies:

``` r
# Install dependencies
install.packages(c("dplyr", "data.table"))

# Install salic from binary executables
install.packages("bla bla") # for Windows
install.packages("bla bla") # for Mac

# Alternatively, install salic from source
install.packages("devtools")
devtools::install_github("southwick-associates/salic")
```

## Usage

See [Introduction to salic](https://southwick-associates.github.io/salic/articles/salic.html) (`vignette("salic")` from the R console).

A template workflow for national/regional dashboards is available at https://github.com/southwick-associates/dashboard-template

### Example: fishing churn rate

Using `rank_sale()`, `make_history()`, `est_part()`, and `format_result()` from `?salic`.

```{r, message=FALSE}
library(dplyr)
library(salic)

# load standardized data
data(cust, lic, sale)

# build a license history
history <- lic %>% 
    filter(type %in% c("fish", "combo")) %>% 
    inner_join(sale, by = "lic_id") %>% 
    rank_sale() %>% 
    make_history(yrs = 2008:2018) %>% 
    left_join(cust, by = "cust_id")

# summarize participation for use in a dashboard
recode_agecat(history) %>%
    filter(!agecat %in% c("0-17", "65+")) %>%
    est_churn() %>%
    format_result(timeframe = "full-year", group = "fish")
```
