---
output: github_document
---
<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/"
)
```

# salic

salic is an R package for preparing and summarizing agency data. In particular, it aims to simplify the production of national/regional dashboards by providing a set of functions that work across state agencies.

## Installation

- First, install R: https://www.r-project.org/
- Then install salic and it's dependencies:

``` r
# Install dependencies
install.packages(c("dplyr", "data.table"))

# Install salic release
f <- "https://github.com/southwick-associates/salic/releases/download/2.0.0/salic_2.0.0"
install.packages(paste0(f, ".zip"), repos = NULL)   # for Windows
install.packages(paste0(f, ".tgz"), repos = NULL)   # for Mac
install.packages(paste0(f, "tar.gz"), repos = NULL) # or from source

# Alternatively, install development version
install.packages("devtools")
devtools::install_github("southwick-associates/salic")
```

## Usage

See [the vignette](https://southwick-associates.github.io/salic/articles/salic.html) for an introduction.

A template workflow for national/regional dashboards is available at https://github.com/southwick-associates/dashboard-template

### Example: fishing participation

Using `rank_sale()`, `make_history()`, and `est_part()`

```{r, message=FALSE}
library(dplyr)
library(salic)

# load data
data(lic, sale)

# build license history
history <- lic %>% 
    filter(type %in% c("fish", "combo")) %>% 
    inner_join(sale, by = "lic_id") %>% 
    rank_sale() %>% 
    make_history(yrs = 2008:2018)

# count total participants
est_part(history)
```
