---
title: "Load Raw License Type Data into sqlite database"
date: '`r strftime(Sys.time(), format = "%B %d, %Y")`'
output: 
    html_document:
        toc: true
        toc_float: true
        code_folding: show
---

```{r setup, message=FALSE}
library(tidyverse)
library(salic)
knitr::opts_chunk$set(comment = NA)
```

**Notes**

In case you want to include some easily accessible final notes

**Guidelines**

- Any cleaning should be minimal (e.g., adding row IDs, dropping empty rows)

- Regarding file paths:
    + Using relative paths to data by going backward (with ../../etc) is a bit awkward, but it will make the code easier to re-run if the path to the SA folder ever changes


## Load

**Guidelines**

I recommend using the package readr functions (read_csv, etc.) because they are fast (for loading large data files) and provide useful ways to work around problems that often arise when loading data (parsing errors, etc.)

```{r}
# path to raw data file
f <- "../../../../../Data-sensitive/Data-Dashboards/__state__/raw-__period__/"

# test import method
read_lines(f, n_max = 3)
read_csv(f, n_max = 5) %>% glimpse() # see readr documentation for non-csv formats

# check line count to ensure correct number of rows after data import
# if this doesn't work, an alternative is to open the file with JujuEdit
salic::count_lines_textfile(f)
```


**Guidelines** 

Running `spec()` is useful to see the decisions readr is making about choosing variable type. Incorrectly specifying type can often lead to problems. Using `col_character()` tends to lead to the least problems, but this is less space-efficient for storing integers or numbers.

```{r}
# check readr column type specification
read_csv(f, n_max = 5) %>% spec() 

# column type specification to use for import (varies by state)
# note: start with output from read_csv() %>% spec()
coltyp <- cols( )

# load full dataset
lic <- read_csv(f, col_types = coltyp, progress = FALSE)

# check for problems
problems(lic)

# get an overview of imported data
glimpse(lic)
```


## Add Row IDs

```{r}
lic$raw_lic_id <- as.integer(row.names(lic))
tail(lic$raw_lic_id)
```


## Write to sqlite

```{r}
db_path <- "../../../../../Data-sensitive/Data-Dashboards/__state__/raw-__period__.sqlite3"
db <- src_sqlite(db_path, create = TRUE)
copy_to(db, lic, temporary = FALSE)
```

```{r}
sessionInfo()
```
