---
title: "Initial Overall Data Check"
date: '`r strftime(Sys.time(), format = "%B %d, %Y")`'
output: 
    html_document:
        toc: true
        toc_float: true
        code_folding: show
---

**Notes**


```{r setup, message=FALSE}
library(tidyverse)
library(salic)
knitr::opts_chunk$set(comment = NA)
```

*Guidelines: If the data from the state has problems, it often shows up in the `summary_sale()`. The number of customers doesn't usually change more than around 10% per year. Big fluctuations might indicate data issues.*


## Load

*Guidelines: only two sales columns are needed (cust_id, year)*

```{r}
f <- "../../../../../Data-sensitive/Data-Dashboards/__state__/raw-__period__.sqlite3"
db_raw <- src_sqlite(f)
src_tbls(db_raw)
```


*Guidelines: Column names will vary by state. Converting to standard names here for `summary_sale()`*

```{r}
sale <- tbl(db_raw, "sale") %>% 
    select(year = TXNL_YEAR, cust_id = TXNL_CUST_ID) %>%
    collect()
glimpse(sale)
```


### Summarize

```{r}
salic::summary_sale(sale)
```


```{r}
sessionInfo()
```
