---
title: "Write Standardized Intermediate Customer Data to Sqlite"
date: '`r strftime(Sys.time(), format = "%B %d, %Y")`'
output: 
    html_document:
        toc: true
        toc_float: true
        code_folding: show
---

**Notes**

**Guidelines**

Saving an intermediate (standardized) database is useful for expediting downstream validation of customer and sales data. This also gets us about halfway toward the final production data, which simplifies the final step of writing production data. 

A select set of variables should be included in the standardized database (see O365 > _Analyst Docs > Data Dictionary...xlsx). It may be worthwhile to include additional variables for certain states, if we expect extra variables to be useful in the dashboard production.


```{r setup, message=FALSE}
library(tidyverse)
library(stringr)
library(salic)
library(lubridate)
knitr::opts_chunk$set(comment = NA)
```

## Load

```{r}
db <- src_sqlite("../../../../../Data-sensitive/Data-Dashboards/__state__/raw-__period__.sqlite3")
src_tbls(db)
```

```{r}
# tbl(db, "cust") %>% glimpse()
cust <- tbl(db, "cust") %>% 
    select(cust_id = CUST_ID, raw_cust_id, sex = CUST_GENDER, dob = CUST_BIRTH_DATE,
           last = CUST_LAST_NAME, first = CUST_FIRST_NAME, state = CUST_STATE,
           zip = CUST_POSTAL_CODE) %>%
    collect()
glimpse(cust)
```


## Prep

**Guidelines**

The code below will need to be modified for state-specific data formats 

### Gender

```{r}
# count(cust, sex)
cust <- cust %>% mutate(
    sex_new = ifelse(sex == "M", 1L, ifelse(sex == "F", 2L, NA))
)
cust %>%
    count(sex_new, sex) %>%
    mutate(pct = round(n / sum(n) * 100, 1))
```

### Date of Birth

**Guidelines**

It's a good idea to look for odd values here, which could be stand-ins for missing values

```{r}
# select(cust, dob) %>% sample_n(5)
cust <- cust %>%
    mutate(
        dob_date = ymd(dob),
        dob_new = as.character(dob_date)
    )
cust %>%
    sample_n(5) %>%
    select(dob_new, dob_date, dob)

# check missing
cust %>%
    filter(is.na(year(dob_date))) %>%
    nrow()

cust %>%
    count(year = year(dob_date)) %>%
    ggplot(aes(year, n)) +
    geom_col() +
    ggtitle("Birth Year customer distribution")
```

### Name

```{r}
# select(cust, first, last) %>% sample_n(5)
cust <- cust %>% mutate(
    last_new = str_trim(last) %>% str_to_lower(),
    first_new = str_trim(first) %>% str_to_lower()
)
cust %>%
    sample_n(5) %>%
    select(first_new, first, last_new, last)

# check missing
cust %>% filter(is.na(first_new) | is.na(last_new)) %>% nrow()
cust %>%
    filter(is.na(first_new) | is.na(last_new)) %>%
    select(first_new, first, last_new, last)
```

### State

```{r}
# select(cust, state) %>% sample_n(5)
data(state_abbreviations)
cust <- cust %>%
    recode_state(state_abbreviations)

cust %>%
    count(state_new, state) %>%
    arrange(desc(n)) %>%
    DT::datatable()

# check missing
# cust %>% filter(is.na(state_new)) %>% nrow()
cust %>%
    filter(is.na(state_new)) %>%
    count(state_new, state) %>%
    arrange(desc(n))
```

### Residency

```{r}
cust <- cust %>% mutate(
    cust_res = ifelse(!is.na(state_new) & state_new == "__state__", 1L, 
                      ifelse(!is.na(state), 0L, NA))
)
cust %>%
    count(cust_res, state_new) %>%
    arrange(desc(cust_res), desc(n)) %>%
    mutate(pct = round(n / sum(n) * 100, 1)) %>%
    DT::datatable()

# check missing
cust %>%
    filter(is.na(cust_res)) %>%
    count(state_new, state)
```

### Zipcode

```{r}
# count(cust, nchar(zip))
cust <- cust %>% mutate(
    # drop non-numeric characters
    zip_clean = str_replace_all(zip, "[^0-9.]", "") %>% str_trim(),  
    zip_new = ifelse(nchar(zip_clean) == 5, zip_clean,
                     # some zips are often zip+4 coded (truncating simplifies things)
                     ifelse(nchar(zip_clean) >= 9, str_sub(zip_clean, end = 5), NA))
)
cust %>%
    sample_n(5) %>%
    select(zip_new, zip)

cust %>%
    count(nchar(zip_new), nchar(zip)) %>%
    mutate(pct = round(n / sum(n) * 100, 1)) %>%
    arrange(desc(n))

# check 9-digit+ truncation
cust %>%
    filter(nchar(zip) >= 9) %>%
    select(zip_new, zip) %>%
    group_by(nchar(zip)) %>%
    sample_n(2)
```

## Finalize

- Renaming "new" columns and dropping old and unnecessary
- Adding a "source" column in case future data is appended

```{r}
cust <- cust %>% select(
    cust_id, sex = sex_new, dob = dob_new, last = last_new, first = first_new,
    state = state_new, cust_res, zip, raw_cust_id
) %>%
    mutate(source_cust = as.integer("__period__"))
glimpse(cust)
```

## Write to Sqlite

```{r}
f <- "../../../../../Data-sensitive/Data-Dashboards/__state__/standard.sqlite3"
db_out <- src_sqlite(f, create = T)
copy_to(db_out, cust, temporary = FALSE)
```

```{r}
sessionInfo()
```
