---
title: "Introduction to salic"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE, message = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(dplyr)
devtools::load_all() # eventually change to library(salic)
```

```{r}
# playing around
library(ggplot2)
data(dashboard)

dashboard %>%
    filter(metric == "participants", segment == "Gender") %>%
    ggplot(aes(year, value, fill = category)) +
    geom_col() +
    facet_wrap(~ group) +
    ggtitle("Hunter vs. Anglers over Time",
            "'all_sports' refer to those who are both hunters and anglers")
```

```{r}
data(history)
x <- history %>%
    label_categories() %>%
    recode_agecat()

x %>%
    filter(age_year >= 25, age_year <= 55, !is.na(R3), !is.na(lapse)) %>%
    group_by(year, R3) %>%
    summarise(churn = mean(lapse)) %>%
    ggplot(aes(year, churn)) +
    geom_col() +
    facet_wrap(~ R3, nrow = 1) +
    ggtitle("Churn rates for different R3 groups for all_sports permission")
```

Processing data for dashboard involves successive steps:

1. **Standardize Data**: First get necessary license data into a standard (anonymized) format well-suited for generalized data processing functions.

2. **Build License History**: Convert into a license history, which tracks license holding at the permission level across years, enabling easy calculation of dashboard metrics.

3. **Calculate Dashboard Metrics**: Apply estimation functions to license history data, which produces the summaries used as input to the dashboard software.

## TODO

- obviously do some outlining/thinking about a good narrative, could potentially use the existing Word documentation to help this process.
- running R code to show sample data, operations,

## Standardizing License Data

These need to be prepared into a standardized format (which is also anonymized to remove any personally-identifiable information).

This process is state-specific since agencies vary in how license data is stored. The standardization is intended to produce a standard format that is well-suited for the generalized presented here. An important part of this step involves anonymizing data by excluding any personally identifiable information from the standard storage format.

### Sample Data

To get started, first load package dependencies:

```r
library(salic)
library(dplyr)
```

Samples of standardized data are provided in salic:

```{r}
data(cust)
glimpse(cust)

data(lic)
glimpse(lic)

data(sale)
glimpse(sale)
```

A function is provided to check that your data fits the specifications (strict for this reason...):

```r
# produces no output (i.e., silent) if no issues detected
check_format(cust, lic, sale)
```

## Building License History

- `rank_sale()` to do this...
    + `join_first_month()` to do that...
    
- `make_lic_history()`
    + `identify_R3()`
    + `identify_lapse()`

Need to also consider the key parameters for this process:

- permission: specifically related to license types include (fish, hunt, combo)
- timeframe: full-year or mid-year
- years

## Calculating Dashboard Metrics

- filter by age
- `label_categories()` & `recode_agecat()`
- `est_part()`, `est_recruit()`, `est_churn()`

### Formatting for final Output
