---
title: "Get ACS Data using Census API with R Package acs"
date: '`r strftime(Sys.time(), format = "%B %d, %Y")`'
output: 
    html_document:
        toc: true
        toc_float: true
        code_folding: show
params:
    state: XX
    state_name: Xxxxxx # Missouri, Georgia, etc.
    lic_yrs: 2007:2016  # years included in license data
    acs_last_yr: 2015
    census_pop1: "~/data2/_CENSUS_ACS/pop_by_state update/st-est00int-01.xls"
    census_pop2: "~/data2/_CENSUS_ACS/pop_by_state update/nst-est2016-01.xlsx"
    test: TRUE  # (TRUE, FALSE) > if TRUE, doesn't write output to sqlite
    db_file: "~/data2/XX_Dashboard/lic.sqlite3"  # sqlite db location (varies by state)
---

```{r setup, message=FALSE}
library(acs)
library(stringr)
library(plyr)
library(tidyverse)
library(readxl)
library(DT)
library(RSQLite)
library(tibble)
source("acs-pop-functions.R")
knitr::opts_chunk$set(comment = NA)
```

## Params

- This script should require minimal updating:
    + acs_last_yr as more data becomes available in package ACS
    + census_pop2 as more data becomes available in US Census. This needs to be downloaded as an Excel file:
    + https://www2.census.gov/programs-surveys/popest/tables/2010-2016/state/totals/

```{r}
# make a vector of years from the params$yrs input
get_range <- function(x) {
    y <- stringr::str_split_fixed(x, ":", 2)
    `:`(y[1], y[2])
}
lic_yrs <- get_range(params$lic_yrs)
```

```{r}
tribble(
    ~param, ~value,
    "state", params$state,
    "state_name", params$state_name,
    "lic_yrs", params$lic_yrs,
    "acs_last_yr", params$acs_last_yr,
    "census_pop1", params$census_pop1,
    "census_pop2", params$census_pop2,
    "test", params$test,
    "db_file", params$db_file
) %>% DT::datatable()
```

## Load Data

### ACS Pop by county by age by sex

```{r}
# check to insure most recent available year is being used (params$acs_last_yr)
# if it shows available tables, then update params$acs_last_yr accordingly
acs.lookup(endyear = params$acs_last_yr+1, table.number = "B01001", span = 5)

# get data
dat <- list()
for (i in 2009:params$acs_last_yr) dat[[i]] <- get_acs(i, params$state) %>% mutate(year = i)
dat <- bind_rows(dat) %>%
    mutate(county = stringr::str_replace(
        county, paste0(" County, ", params$state_name), "")
    )
# count(dat, year)
# count(dat, county) %>% View()
glimpse(dat)
```

### US Census Pop by state

```{r}
pop_state <- list(
    get_state_pop_old(params$census_pop1, 2000:2009),
    get_state_pop(params$census_pop2, 2010:2016)
)
pop_state <- bind_rows(pop_state) %>%
    mutate(year = as.numeric(year)) %>%
    filter(state == params$state_name)
data.frame(pop_state)
```

```{r}
# Example Code if years more recent than last needed
# estimate 2017 - assume linear trend
# model <- lm(pop_state ~ year, data = pop_state)
# est <- predict(model, data.frame(year = lic_yrs))
# pop_state <- add_row(pop_state, state = "Georgia", year = 2017, pop_state = est[8])
# 
# data.frame(pop_state)
```

### County fips codes

```{r}
data("fips.county", package = "acs")
fips_county <- filter(fips.county, State == params$state) %>%
    mutate(county = stringr::str_replace(County.Name, " County", "")) %>%
    select(fips = County.ANSI, county)
rm(fips.county)
# DT::datatable(fips_county)
```

## County Coverage

- n should be the same for all counties

```{r}
full_join(fips_county, dat, by = "county") %>%
    count(fips, county) %>%
    DT::datatable()
```

### County Adjustments

- A Missouri example, probably not needed for most states

```{r}
dat_in <- dat

# fix naming for St Louis City
dat <- mutate(dat, county = str_replace(county, ", Missouri", ""))
full_join(fips_county, dat, by = "county") %>%
    count(fips, county) %>%
    DT::datatable()

# combine St Louis city and St Louis into a single county
# x <- filter(dat, county %in% c("St. Louis city", "St. Louis"))
# dat <- anti_join(dat, x)
# nrow(x) + nrow(dat) == nrow(dat_in)
# 
# st_louis <- group_by(x, var, year) %>%
#     summarise(value = sum(value)) %>%
#     mutate(county = "St. Louis")
# dat <- bind_rows(dat, st_louis)
# 
# # check > st louis city is dropped
# right_join(fips_county, dat, by = "county") %>%
#     count(fips, county) %>%
#     DT::datatable()
```


## Prep ACS

### Tidy column names

```{r, warning=FALSE}
# a "Too few values" warning is expected since sex & age aren't populated for all var
# it's probably not an indication of a problem
dat_in <- dat

dat <- mutate(dat_in, cat = str_replace(var, "Sex by Age: ", "")) %>%
    tidyr::separate(cat, c("sex", "age"), sep = ": ") %>%
    mutate(
        sex = ifelse(sex == "Total:", "All", str_replace(sex, ":", "")),
        age = ifelse(is.na(age), "All", age)
    )
count(dat, sex, age, var) %>% DT::datatable()
```

### Totals & Pct Change

```{r}
check_pop_totals(dat) # all diff should be zero
pct_change_sex(dat) %>% select(-value) %>% spread(sex, pct_change) %>% data.frame()
pct_change_age(dat) %>% select(-value) %>% spread(year, pct_change) %>% data.frame()
```

### Drop "All" rows

- totals can easily be calculated by grouping, so no need to keep these
- The overall county totals for the most recent year will be kept temporarily for the summary at the end of the script

```{r}
dat_total <- filter(dat, year == params$acs_last_yr, sex == "All", age == "All")
dat <- filter(dat, sex != "All", age != "All") %>%
    select(-var)
glimpse(dat)
```

## Adjust with Census

### Match Totals by Year

- This insures the county totals match the most recent estimates available from the Census
- The adjustments should be pretty small (maybe a couple percentage points)

```{r}
# estimate necessary adjustment
adjust <- group_by(dat, year) %>%
    summarise(value = sum(value)) %>%
    left_join(pop_state, by = "year") %>%
    mutate(adjust = pop_state / value)

# apply adjustment across all detailed estimates
dat <- select(adjust, year, adjust) %>%
    full_join(dat, by = "year") %>%
    mutate(value = value * adjust) %>%
    select(-adjust) %>%
    ungroup()

# check
group_by(dat, year) %>%
    summarise(new_value = sum(value)) %>%
    left_join(adjust)
```

### Fill missing years

- for missing years (pre 2009 & post most recent available years)
- populate using population by state

```{r}
pct <- list()  # distributions
est <- list()  # estimated populations

# pre 2009 > base on 2009 distributions
pct[[1]] <- filter(dat, year == 2009) %>%
    left_join(pop_state, by = "year") %>%
    mutate(pct = value / pop_state) %>%
    select(-year, -value, -pop_state)

est[[1]] <- filter(pop_state, year %in% lic_yrs[1]:2008) %>%
    left_join(pct[[1]], by = "state") %>%
    mutate(value = pop_state * pct) %>%
    select(-pop_state, -pct, -state)

# post acs_last_yr > based on acs_last_yr
pct[[2]] <- filter(dat, year == params$acs_last_yr) %>%
    left_join(pop_state, by = "year") %>%
    mutate(pct = value / pop_state) %>%
    select(-year, -value, -pop_state)

yr1 <- params$acs_last_yr+1
est[[2]] <- filter(pop_state, year %in% yr1:lic_yrs[length(lic_yrs)]) %>%
    left_join(pct[[2]], by = "state") %>%
    mutate(value = pop_state * pct) %>%
    select(-pop_state, -pct, -state)

# combine & check
dat <- bind_rows(est[[1]], dat, est[[2]])

group_by(dat, year) %>%
    summarise(value = sum(value)) %>%
    left_join(pop_state)
```

## Final Prep

### Recode variables

- To match the coding format of the state license data
    
```{r}
dat <- dat %>% mutate(
    sex_old = sex,
    sex_label = factor(sex, c("Male", "Female")),
    sex = as.numeric(sex_label)
)
count(dat, sex, sex_label, sex_old) %>% data.frame()
```

```{r}
age_map <- tribble(
    ~lic_age, ~acs_age,
    1,"Under 5 years",
    1,"5 to 9 years",
    1,"10 to 14 years",
    1,"15 to 17 years",
    2,"18 and 19 years",
    2,"20 years",
    2,"21 years",
    2,"22 to 24 years",
    3,"25 to 29 years",
    3,"30 to 34 years",
    4,"35 to 39 years",
    4,"40 to 44 years",
    5,"45 to 49 years",
    5,"50 to 54 years",
    6,"55 to 59 years",
    6,"60 and 61 years",
    6,"62 to 64 years",
    7,"65 and 66 years",
    7,"67 to 69 years",
    7,"70 to 74 years",
    7,"75 to 79 years",
    7,"80 to 84 years",
    7,"85 years and over"
)
dat_tmp <- dat
dat <- dat %>% mutate(
    age_acs = age,
    age_acs_ordered = factor(age, levels = age_map$acs_age),
    age = plyr::mapvalues(age, age_map$acs_age, age_map$lic_age) %>% as.integer(),
    age_label = factor_age(age)
)
count(dat, age, age_label, age_acs_ordered, age_acs) %>% data.frame()
```

### Collapse over 7 age categories

- this is necessary since the ACS tracks many more age categories than we will track in the dashboard

```{r}
dat <- group_by(dat, year, county, sex, age) %>%
    summarise(value = sum(value)) %>%
    ungroup()
glimpse(dat)
```

### Join Fips

```{r}
dat <- mutate(dat, year = as.integer(year), sex = as.integer(sex)) %>%
    left_join(fips_county, by = "county")
```

### Pop by County

- Summarizes yearly values for total population by county
- Check some columns for consistency
    + pct_sd - will probably be less than 10% or so
    + n values (year, total_county) should be the same for all counties
    + pct_diff (last year versus mean) - probably will be within 10% or so

- Could also compare to wikipedia: For example,
    + https://en.wikipedia.org/wiki/List_of_counties_in_Missouri

```{r}
x <- count(dat, county) %>% rename(n_total_county = n) 
group_by(dat, county, year) %>%
    summarise(value = sum(value)) %>%
    summarise(min = round(min(value), 0), mean = round(mean(value), 0), 
              max = round(max(value), 0),  sd = round(sd(value), 0), 
              pct_sd = round(sd / mean * 100, 3), n_year = n()) %>% 
    left_join(x) %>%
    left_join(select(dat_total, county, acs_last_year = value)) %>%
    mutate(pct_diff = round(abs(acs_last_year - mean) / mean * 100, 2)) %>%
    DT::datatable()
```

## Write to SQLite

```{r}
pop_county <- select(dat, year, fips, sex, age, pop = value) %>%
    data.frame()
glimpse(pop_county)
glimpse(fips_county)
```

```{r}
# write to sqlite database
if (!params$test) {
    con <- dbConnect(SQLite(), dbname = params$db_file)
    
    # write tables (overwrite if this table already exists)
    if ("pop_county" %in% dbListTables(con)) dbRemoveTable(con, "pop_county")
    dbWriteTable(con, "pop_county", pop_county)
    
    if ("fips_county" %in% dbListTables(con)) dbRemoveTable(con, "fips_county")
    dbWriteTable(con, "fips_county", fips_county)
    
    dbDisconnect(con)
}
```

```{r}
sessionInfo()
```

